Corporate Security Solutions for BYOD:\\ A Novel User-Centric and Self-Adaptive System

### 2015-06-16 MUSES_vs_BYOD_solutions.tex
un comentario
Removed all slashes in the text (changed by 'ors' and 'ands'). Closes #38

The rest of the paper is organized as follows. First, some background situation about enterprise security is presented in Section \ref{sec:preliminaryconcepts}, explaining how BYOD philosophy affects it. Then, the main tools/solutions are detailed in Section \ref{sec:toolsreview}, even if they are still in development.  
Then, in Section \ref{sec:muses}, the Multiplatform Usable Endpoint Security System is presented (architecture + soft computing techniques). Its advantages and benefits in comparison with the other solutions are commented in Section \ref{sec:comparison}.
Finally, the reached conclusions are discussed in Section \ref{sec:conclusions}.

@@

One of the tasks of a Chief Security Officer (CSO) is to elaborate an efficient set of Information Security Policies (ISPs), for controlling a certain, and already known, structure \cite{Opp_Security11}. This means that, until the appearance of BYOD, enterprises used to follow static security policies devoted to control an environment where the company assets and the devices were purchased and maintained by the company. Now that corporate networks are becoming dynamic for being adapted to this BYOD philosophy, there is an additional risk because the devices that the employees use are not always company-owned. This means that the CSOs have to find the balance between having a fast response to any user action that might cause harm (in terms of money loss because of a security incident), and trying to avoid monitoring the users in a way that is against privacy. A needed security policy, or in this case, an ISP should deal with the way of protecting a certain organization's information against a security breach. Though there are standards, such as the ISO27002 or the Security Forum's Standard of Good Practice\footnote{\url{https://www.securityforum.org}}, and many guidelines \cite{SecPol09}, an ISP should be adapted depending on the characteristics of the community/organisation that they are built for. In this sense, MUSES stands up for a self-adaptation of the security policies of a company.

@@

The other issue to cope with is the elaboration of a good ISP, understandable for every user of the company, and more importantly, non-intrusive for him. A lot of researchers have studied the natural tendency of employees to comply with the ISP \cite{SecPolComp10,SecPolComp12,SecPolComp14}, reaching conclusions such as the employees compliance with the security policies increases educating/training them in information security awareness  \cite{SecPolComp09}, and decreases applying too much sanctions when a misuse or abuse occurs \cite{SecPolPenalty09}. Thus, in this paper, for each tool in Section \ref{sec:toolsreview} it is specified if the developers have taken into account the construction of ISPs, or even if they give some guidelines for building them.

@@

This system will be strongly related to the Google Play market, as the applications will be `categorized'. Thus, Android Work will provide the CSOs with a tool to define which apps would be allowed for being installed by the users and considered as corporate-related applications. Several enterprise-security aimed applications will be offered in the market, so the IT would decide which ones will be installed (even as a bundle) in the employees' devices, and also, they can manage the updates of these apps, in order to ensure that all the employees are up to date. For instance, one app could manage the creation of personal/work profiles, so the user just could access corporate assets after login in into the app. Moreover, the enterprise IT department could define specific security policies on these apps, in addition to the inherent protection that the container provides. However, the specifications of this tool \cite{AndroidWork_tool} do not include the ``self-adaptation'' as a feature, as the MUSES system does.

@@

\begin{itemize}
\item It offers a central deployment, administration and security control of all mobile devices regardless of their OS. %FERGU: securing como palabra no. security control mejor?
%Cambiado
\item It manages and enforces security policies `over the air' for both corporate and employee-owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe/selective wipe corporate data email/contacts on the device if lost or stolen.
\end{itemize}

@@

The designed MUSES architecture is shown in Figure \ref{fig:architecture}. It is a \textit{client/server} approach in which the \textit{client} application can be installed in every user's mobile or portable device, independently of the platform. On the one hand, MUSES client has been developed for Android, to be used in Android smartphones and tablets, as well as for Windows, so it can be used in Windows devices and laptops, and finally, iOS client will be also developed. On the other hand, MUSES server has been developed in Java, and works with an Apache Tomcat server. The \textit{server} side can be installed in the corporate server. Both sides are connected through a secure channel (using HTTPS) over the Internet. In that figure just the high-level components in each part are shown, along with the way the information flows in the system. %FERGU: esto no se si es de muses y posiblemente esté por terminar, pero quitad las comillas raras

@@

All flows of information in MUSES are encrypted with AES with 128-bit key in GCM mode \cite{viega2005use}, following the recommendations fo the Fact Sheet on Suite B Cryptography \cite{salter2012suite}. Also, the MUSES server authenticates to the MUSES client by means of a self-signed certificate. The authentication of the users is based on Spring Security, which is a Java/Java EE framework that provides authentication, authorization and other security features for enterprise applications. And finally, MUSES makes use of the fact that current operating systems for mobile devices provide encrypted data containers by default, like the feature used in Android Work.

@@

On the other hand the \textit{Non MUSES Aware App} is that which MUSES cannot directly interact with. Usually it can be accessed through the operating system (OS). So the type of information to gather from them and also the type of actions to do over them vary depending on the OS. In most of the cases MUSES could just detect/recognize very specific events, and also do very limited actions to prevent or block them (maybe showing a screen on the top until something is accomplished by the user). This is due to the usual denial or limitation of interactions with other applications that the OSs have. %FERGU: forbiddance parece que existe, pero el corrector me dice que no. Prohibition mejor?
% Yo diría que es forbidding...

@@

As previously stated, one of the main features of this system is the self-adaptation (to the user and context), of the set of Corporate Security Rules (specification of the ISPs), that it is be able to perform. 
To this end, the \textit{MusKRS} module is run asynchronously in the server (once a day or every X hours) and it is in charge of analysing all the gathered information (events, context, user-related data), and adapting/refining the security rules. The refinement (improvement)  has as aim to better deal with these events, optimizing the current set of rules (removing redundant or non-useful rules, for instance), and even trying to predict future threats due to the user's behaviour.

@@

The following sections describe these processes: first focusing on DM techniques to be used both automatically by the MusKRS, and as a kind of decision-aid/monitoring tool for the CSO; second, the CI techniques are explained, mainly focusing in Evolutionary Computation (EC) \cite{EAs_Back96} approaches, since these methods perform very well, and have been widely used in security-based environments for solving security issues, such as the intrusion detection \cite{GA_intrusion-majeed}, the design and evaluation of security protocols \cite{GP_intrusion-lu,GA_networksecurity-zarza}, or the optimisation of different aspects related with security: IT security costs \cite{EAs_securitycosts-kirta}, and cryptographic protocols \cite{GA_cryptographicprotocols-zarza2}, among others.

@@

\item \textit{Clustering} \cite{Clustering_Jain99}: 
The aim of this method is grouping the patterns considering some similarity criteria, in order to manage them as a set. This could be used for providing data visualisation mechanisms, in order to make it easier to interpret the data interaction and the distribution in clusters with respect to the different properties/features of the patterns.

\item \textit{Feature Selection} \cite{FeatureSelection_Guyon03}: 
It consists on extract the most important features/variables from the data. This could be useful if we want to discard non-key features, which could be interesting in order to reduce the database weight, for improving the performance of other techniques (such as classification or clustering), and even to improve the performance of the whole system, since less information would be gathered and transmitted.

@@

MUSES  uses different EC approaches, initially, two in the DM/ML part of the process, and three in the rule-refinement/adaptation phase. They are based on Genetic Programming (GP) \cite{GP_Koza92}, and Genetic Algorithms (GAs) \cite{GAs_Goldberg89}.

The first evolutionary-based approach is a \textit{GP-based classification} method. This is useful for two main reasons: first, in order to deal with the data class imbalance \cite{imbalance_techniques_02}, very common in classification problems in real systems (with real data); and second, to better manage categorical (non-numeric) data, since most of the features/variables and information gathered from the events take these kind of values.

@@

The second approach is a Feature Selection process, conducted using a GA as a meta-algorithm to test all the possible combinations of pattern features/variables in the classification. Thus, the algorithm can optimize the group of features to be considered in the classification task, improving both the computational time of this task and also the obtained accuracy of the method.

@@

Other useful information to consider in the refinement/ adaptation process are:

@@

\item Risk information extracted from the user profile (user trust value), e.g. ``Did the user received a lot of `denies'/`allows' before?'', i.e. ``Is the user trustworthy?''. In case the user is not, more restrictive rules can be created, otherwise the corresponding rules could be `relaxed' for that user.

@@

\item \textit{GP rule inference} method, which generates/creates new rules in order to `cover' those situations non contemplated in the current set of rules. Thus, for instance, a new rule could be created in order to deal with the patterns to which the classifier could not assign a class.
The generation can be done considering the so-called \textit{dictionary}, i.e. a set of terms corresponding to all the possible inputs and actions in the system, which are antecedents (conditions) and consequents in the security rules to be inferred.
The evaluation of these rules can be done considering the stored log information concerning the parameters along with the actions/decisions made in every component in the system. Thus, it is possible to `simulate' the whole  system behaviour when the new rule is included and get a value of its performance.

@@

Another comparison would be that the existing products are mostly policy-based, while MUSES makes its decisions not only considering policies, but also based on the terminals/users context (connection properties, or trust value of the device, for instance), to really understand the real risk that comes with a specific action.

@@

With regard to the scientific contribution of the system, one of the main differences with respect to previous works is the consideration of security threats `brought' by the user's behaviour inside the system, i.e. through interaction/events, rather than more general and external threats. Moreover, the techniques to be used in MUSES  works with real data (in a real system), as a difference to several research works (they consider simulated or artificial data).

@@

MUSES has been compared with other existing solutions/tools with a similar aim: to ensure the protection and privacy of sensitive information in the enterprise when it is accessed by the employees using their own devices (BYOD philosophy).
The main advantages of MUSES with respect to the other solutions have been discussed, also remarking its advances beyond the state of the art that define these tools.

>>>>

The rest of the paper is organized as follows. First, some background situation about enterprise security is presented in Section \ref{sec:preliminaryconcepts}, explaining how BYOD philosophy affects it. Then, the main solutions or tools are detailed in Section \ref{sec:toolsreview}, even if they are still in development.  
Then, in Section \ref{sec:muses}, the Multiplatform Usable Endpoint Security System is presented (architecture + soft computing techniques). Its advantages and benefits in comparison with the other solutions are commented in Section \ref{sec:comparison}.
Finally, the reached conclusions are discussed in Section \ref{sec:conclusions}.

@@

One of the tasks of a Chief Security Officer (CSO) is to elaborate an efficient set of Information Security Policies (ISPs), for controlling a certain, and already known, structure \cite{Opp_Security11}. This means that, until the appearance of BYOD, enterprises used to follow static security policies devoted to control an environment where the company assets and the devices were purchased and maintained by the company. Now that corporate networks are becoming dynamic for being adapted to this BYOD philosophy, there is an additional risk because the devices that the employees use are not always company-owned. This means that the CSOs have to find the balance between having a fast response to any user action that might cause harm (in terms of money loss because of a security incident), and trying to avoid monitoring the users in a way that is against privacy. A needed security policy, or in this case, an ISP should deal with the way of protecting a certain organization's information against a security breach. Though there are standards, such as the ISO27002 or the Security Forum's Standard of Good Practice\footnote{\url{https://www.securityforum.org}}, and many guidelines \cite{SecPol09}, an ISP should be adapted depending on the characteristics of the community or organisation that they are built for. In this sense, MUSES stands up for a self-adaptation of the security policies of a company.

@@

The other issue to cope with is the elaboration of a good ISP, understandable for every user of the company, and more importantly, non-intrusive for him. A lot of researchers have studied the natural tendency of employees to comply with the ISP \cite{SecPolComp10,SecPolComp12,SecPolComp14}, reaching conclusions such as the employees compliance with the security policies increases educating or training them in information security awareness  \cite{SecPolComp09}, and decreases applying too much sanctions when a misuse or abuse occurs \cite{SecPolPenalty09}. Thus, in this paper, for each tool in Section \ref{sec:toolsreview} it is specified if the developers have taken into account the construction of ISPs, or even if they give some guidelines for building them.

@@

This system will be strongly related to the Google Play market, as the applications will be `categorized'. Thus, Android Work will provide the CSOs with a tool to define which apps would be allowed for being installed by the users and considered as corporate-related applications. Several enterprise-security aimed applications will be offered in the market, so the IT would decide which ones will be installed (even as a bundle) in the employees' devices, and also, they can manage the updates of these apps, in order to ensure that all the employees are up to date. For instance, one app could manage the creation of personal and work profiles, so the user just could access corporate assets after login in into the app. Moreover, the enterprise IT department could define specific security policies on these apps, in addition to the inherent protection that the container provides. However, the specifications of this tool \cite{AndroidWork_tool} do not include the ``self-adaptation'' as a feature, as the MUSES system does.

@@

\begin{itemize}
\item It offers a central deployment, administration and security control of all mobile devices regardless of their OS. %FERGU: securing como palabra no. security control mejor?
%Cambiado
\item It manages and enforces security policies `over the air' for both corporate and employee-owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe (even selectively) corporate data email and contacts on the device if lost or stolen.
\end{itemize}

@@

The designed MUSES architecture is shown in Figure \ref{fig:architecture}. % [Pedro] creo que la frase "suena mejor" sin "The designed" (pero es una apreciación subjetiva mía).
It is a \textit{client/server} approach in which the \textit{client} application can be installed in every user's mobile or portable device, independently of the platform. On the one hand, MUSES client has been developed for Android, to be used in Android smartphones and tablets, as well as for Windows, so it can be used in Windows devices and laptops, and finally, iOS client will be also developed. % [Pedro] ¿qué tal decir: "MUSES client is available up to date to be used in Android devices as well as in Windows mobile devices and PCs; the iOS client is under development."
On the other hand, MUSES server has been developed in Java, and works with an Apache Tomcat server. The \textit{server} side can be installed in the corporate server.  % [Pedro] y respecto al servidor: "As far as the server is concerned, it has been developed using Java to be used with an Apache Tomcat in the corporate server."
Both sides are connected through a secure channel (using HTTPS) over the Internet. In that figure just the high-level components in each part are shown, along with the way the information flows in the system. %FERGU: esto no se si es de muses y posiblemente esté por terminar, pero quitad las comillas raras

@@

All flows of information in MUSES are encrypted with AES with 128-bit key in GCM mode \cite{viega2005use}, following the recommendations fo the Fact Sheet on Suite B Cryptography \cite{salter2012suite}. Also, the MUSES server authenticates to the MUSES client by means of a self-signed certificate. The authentication of the users is based on Spring Security, which is a Java EE framework that provides authentication, authorization and other security features for enterprise applications. And finally, MUSES makes use of the fact that current operating systems for mobile devices provide encrypted data containers by default, like the feature used in Android Work.

@@

On the other hand the \textit{Non MUSES Aware App} is that which MUSES cannot directly interact with. Usually it can be accessed through the operating system (OS). So the type of information to gather from them and also the type of actions to do over them vary depending on the OS. In most of the cases MUSES could just detect or recognize very specific events, and also do very limited actions to prevent or block them (maybe showing a screen on the top until something is accomplished by the user). This is due to the usual denial or limitation of interactions with other applications that the OSs have. %FERGU: forbiddance parece que existe, pero el corrector me dice que no. Prohibition mejor?
% Yo diría que es forbidding...

@@

As previously stated, one of the main features of this system is the self-adaptation (to the user and context), of the set of Corporate Security Rules (specification of the ISPs), that it is be able to perform. 
To this end, the \textit{MusKRS} module is run asynchronously in the server (once a day or every X hours) and it is in charge of analysing all the gathered information (events, context, user-related data), and adapting or refining the security rules. The refinement (improvement)  has as aim to better deal with these events, optimizing the current set of rules (removing redundant or non-useful rules, for instance), and even trying to predict future threats due to the user's behaviour.

@@

The following sections describe these processes: first focusing on DM techniques to be used both automatically by the MusKRS, and as a kind of decision-aid and monitoring tool for the CSO; second, the CI techniques are explained, mainly focusing in Evolutionary Computation (EC) \cite{EAs_Back96} approaches, since these methods perform very well, and have been widely used in security-based environments for solving security issues, such as the intrusion detection \cite{GA_intrusion-majeed}, the design and evaluation of security protocols \cite{GP_intrusion-lu,GA_networksecurity-zarza}, or the optimisation of different aspects related with security: IT security costs \cite{EAs_securitycosts-kirta}, and cryptographic protocols \cite{GA_cryptographicprotocols-zarza2}, among others.

@@

\item \textit{Clustering} \cite{Clustering_Jain99}: 
The aim of this method is grouping the patterns considering some similarity criteria, in order to manage them as a set. This could be used for providing data visualisation mechanisms, in order to make it easier to interpret the data interaction and the distribution in clusters with respect to the different properties (features) of the patterns.

\item \textit{Feature Selection} \cite{FeatureSelection_Guyon03}: 
It consists on extract the most important features (variables) from the data. This could be useful if we want to discard non-key features, which could be interesting in order to reduce the database weight, for improving the performance of other techniques (such as classification or clustering), and even to improve the performance of the whole system, since less information would be gathered and transmitted.

@@

MUSES  uses different EC approaches, initially, two in the DM/ML part of the process, and three in the rule-refinement and adaptation phase. They are based on Genetic Programming (GP) \cite{GP_Koza92}, and Genetic Algorithms (GAs) \cite{GAs_Goldberg89}.

The first evolutionary-based approach is a \textit{GP-based classification} method. This is useful for two main reasons: first, in order to deal with the data class imbalance \cite{imbalance_techniques_02}, very common in classification problems in real systems (with real data); and second, to better manage categorical (non-numeric) data, since most of the features (variables) and information gathered from the events take these kind of values.

@@

The second approach is a Feature Selection process, conducted using a GA as a meta-algorithm to test all the possible combinations of pattern features (variables) in the classification. Thus, the algorithm can optimize the group of features to be considered in the classification task, improving both the computational time of this task and also the obtained accuracy of the method.

@@

Other useful information to consider in the refinement and adaptation process are:

@@

\item Risk information extracted from the user profile (user trust value), e.g. ``Did the user received a lot of `denies' or `allows' before?'', i.e. ``Is the user trustworthy?''. In case the user is not, more restrictive rules can be created, otherwise the corresponding rules could be `relaxed' for that user.

@@

\item \textit{GP rule inference} method, which generates new rules in order to `cover' those situations non contemplated in the current set of rules. Thus, for instance, a new rule could be created in order to deal with the patterns to which the classifier could not assign a class.
The generation can be done considering the so-called \textit{dictionary}, i.e. a set of terms corresponding to all the possible inputs and actions in the system, which are antecedents (conditions) and consequents in the security rules to be inferred.
The evaluation of these rules can be done considering the stored log information concerning the parameters along with the actions and decisions made in every component in the system. Thus, it is possible to `simulate' the whole  system behaviour when the new rule is included and get a value of its performance.

@@

Another comparison would be that the existing products are mostly policy-based, while MUSES makes its decisions not only considering policies, but also based on the terminals and users context (connection properties, or trust value of the device, for instance), to really understand the real risk that comes with a specific action.

@@

With regard to the scientific contribution of the system, one of the main differences with respect to previous works is the consideration of security threats `brought' by the user's behaviour inside the system, i.e. through interaction or events, rather than more general and external threats. Moreover, the techniques to be used in MUSES  works with real data (in a real system), as a difference to several research works (they consider simulated or artificial data).

@@

MUSES has been compared with other existing solutions with a similar aim: to ensure the protection and privacy of sensitive information in the enterprise when it is accessed by the employees using their own devices (BYOD philosophy).
The main advantages of MUSES with respect to the other solutions have been discussed, also remarking its advances beyond the state of the art that define these tools.
